---
title: "Cointegration"
author: "Tyler J. Brough"
date: "12/3/2016"
output:
  html_document: default
  pdf_document: default
---

# Introduction: Cointegration Notes

These notes are based on material from several sources:

* Chapter 22 of Greene
* Chapter 19 of Kennedy
* Chapter 11 in Cochrane
* Chapter 8 in Brooks


# A Monte Carlo Study to Demonstrate Spurious Regression

## Spurious Regression

Consider the following two ___independent___ unit-root nonstationary processes:

$$
\begin{align}
y_{1,t} &= y_{1,t-1} + u_{t} \\
y_{2,t} &= y_{2,t-1} + v_{t} 
\end{align}
$$

with $E(u_{t}, v_{s}) = 0$ for every $t,s$. 

<br>

A regression on one of these series on the other, such as:

$$
y_{1,t} = \alpha + \beta y_{2,t} + \varepsilon_{t}
$$

_NB:_ we expect the $t$-statistic to fail to reject the null $H_{0}: \beta = 0$, give their independence.


## Monte Carlo Simulation

But if either or both (as in this case!) of $y_{1,t}$ or $y_{2,t}$ is a unit-root process this regression violates the
assumptions of the CLRM and will result in spurious results. 

<br>

To illustrate this, consider the following Monte Carlo simulation:

1. Simulate an articifial dataset consisting of 500 observations from the processes given above.
2. Regress $y_{1,t}$ on $y_{2,t}$ and store the resulting $R^{2}$ and $t$-ratio.
3. Repeat this process for $i = 1, 2, \cdots, M$ with $M = 10,000$.
4. Plot histograms of the $M$ $R^{2}$'s and $t$-ratios. 


## The R Code

```{r message=FALSE, warnings=FALSE}
M <- 10000
T <- 500

rsqrd <- rep(0, M)
tstat <- rep(0, M)

for(i in 1:M)
{
	y <- cumsum(rnorm(T))
	x <- cumsum(rnorm(T))
	fit <- lm(y ~ x)
	rsqrd[i] <- summary(fit)$r.squared
	tstat[i] <- summary(fit)$coefficients[2,3]
}

hist(rsqrd, breaks = 100, col="blue", main = "Histogram of Simulated R^{2}")
hist(tstat, breaks = 100, col="blue", main = "Histogram of Simulated t-ratio")
```


## Cointegration $\phantom{Page 1 of ECN7320 Notes Dated 03/01/2011}$

Cointegration is a generalization of unit roots to vector systems. 

__Definition:__ Suppose that two time series are each integrated, i.e. have unit roots, and hence have the following
moving average representations:

$$
\begin{align}
(1 - L)y_{t} &= a(L)\delta_{t} \\
(1 - L)w_{t} &= b(L)v_{t} 
\end{align}
$$


## $\phantom{Page 2}$

In general linear combinations of $y$ and $w$ also have unit roots. 

<br>

If however, there is some linear combination, say

$$
y_{t} = \alpha w_{t}
$$

that is stationary, $y_{t}$ and $w_{t}$ are said to be ___cointegrated___, and $[1 - \alpha]$ is their ___cointegrating
vector___.


## Some Plausible Examples $\phantom{These plots and data inserted. Not in original notes.}$

Consider the daily closing log-prices for Coke and Pepsi for 2015, shown below.

```{r message=FALSE, warnings=FALSE, code=readLines("plot-coke-pepsi.R")}
```

## 

Consider the bid and ask prices for GOOG for a few hours from a single trading day.

```{r message=FALSE, warnings=FALSE, code=readLines("plot-goog-bid-ask.R")}
```


